
#Exam 1
## By Ledin Menjivar

```{r}
load('turtles.rda')
head(turtles)
```

### Question 1
**1**
```{r}
dim(turtles)

```
The dim() function in R returns the dimension of our data frame in rows by columns. In this case it returns $48$ rows and $4$ columns. 

**2**
```{r}
str(turtles)
```
The str() function returns the structure of the data frame and tells us whats in the data. In this case we see that we have a column with factors and three columns which have integer values.We also see that we have 48 observations which are our rows and 4 variables which are the columns. 

**3**
```{r}
summary(turtles)
```

The summary() function gives a summary of basic statistic with in the data frame. If the data is not a numeric value the it may also include counts for whats in the data.

**4**
```{r}
colnames(turtles)
```

The colnames() function returns the names of each column with in our data.

**5**
```{r}
View(turtles)
```

The view function returns a different window where you can view your data as a table. 

**6**
```{r}
head(turtles)
```
The head() function displays the first 6 rows by default with in our data. The number of rows to display can also be specified. 

**7**
```{r}
tail(turtles)
```

The tail() function will display by default the last 6 rows of our data. The amount of rows can also be specified.

**8**
```{r}
plot(turtles)
```

The plot() function plots one vector vs another and plots the points in a chart. 

###Question 2
**(a)**
```{r}
X1 = turtles$Length
X2 = turtles$Width
X3 = turtles$Height
X = cbind(X1,X2,X3)
```
**(b)**
The first three lines of the R code assigns three different variables to each of the three columns creating three different vectors. 

**(c)**
The last line of code combines the three different vectors by columns. The dimensions of the vectors must also be of the same length. 

### Question 3
**(a)**
```{r}
library(mvnormtest)
 mshapiro.test(t(X))
```

**(b)**
This code chunk tests for normality. And it uses the Shapiro-Wilk test to check for multivariate normality. The function mshapiro.test() must take in a data frame of numeric values to perform the test.

```{r, echo=FALSE}
shapirovalue <- mshapiro.test(t(X))
pValue <- shapirovalue$p.value
```
**(c)**
Using this test for normality we have p-value = $`r pValue`$. This indicates that one or more variables are not normaly distrivuted, therefore multivariate normality does not hold.

**(d)**
Yes. This t() is a function which will transpose our data frame. This is necessary so that our observations are in a column and the variables of those observations are in rows. This is necessary for the mshapiro.test() function to work.

### Question 4
**(a)**
```{r}
X.SM = apply(X,MARGIN = 2, FUN = mean)
X.SM
```

**(b)**
This code chunk will apply the mean function and return the mean of each column in X. 

### Question 5
**(a)**
```{r}
M.VC = var(X)
M.VC
```

```{r}
M.VC = cov(X)
M.VC
```

```{r}
det(M.VC)
```

```{r}
M.VC = cor(X)
M.VC
```

**(b)**
The first chunk of code calculates the variance of the data. 
The second chunk of code calculates the covariance of our data. 

**(c)**
The third chunk of code calculates the determinat of the data. 
The determinant of our data is $12790.67$. Which is a positive value greater than $0$ and tells us we have a nonsingular matrix. Therefore we can proceed with our multivariate statistical analysis. 

**(d)**
The fourth code chunk calculates the correlation matrix of our data.
In the first and second code chunks we have the same matrix. For the third chunk the correlation matrix, we have positive values with $1$ on our diagonals which indicate a positive relationship between variables. 

### Question 6
```{r}
library(biotools)
data(turtles)
factor(turtles[,1])
boxM(turtles[,-1],turtles[,1])
```

**(a)**
The code chunk use the Box M Test which tests the equality of matrices between groups. It will specifically test for equality between variance and covariance matrices. 

**(b)**
We have a P-value of $0.0005141$ and a chi squared value of $24.037$ which indicate we have normality in our data and equal variance and covariance. 

### Question 7
**(a)**
```{r}
#install.packages("ICSNP")
library(ICSNP)
hypvedatact = c(149,100,53)
test = HotellingsT2(X,mu=hypvedatact,test="f")
print(test)
```

**(b)**
The code uses a one sample T2 test and especifically is the two group (dependent) variable mean difference test. The test also uses the F-Distribution for the Hotellings T2 test. Running the code we get $T.2 = 251.2$ and $p-value = 2.2*e^{-16}$. This shows that our mean difference is statistically significant in our data. 

### Question 8

**(a)**
```{r}
#install.packages("ICSNP")
library(ICSNP)
library(dplyr)
HotellingsT2(filter(turtles,Gender=="Male")[,2:4],filter(turtles,Gender=="Female")[,2:4])
```

**(b)**
This code is the Hotelling's two sample T2-test. Especifically is the Two Independent Group Mean Difference test. We get the values $T2 = 21.284$ and $p-value = 1.141*e^{-08}$. Thus we can deduce that the male and female turtles are positively cor-related.











